R tools for accessing research literature for text mining

### 2019-08-23 manuscript.Rmd
fix #7 clarify scanned pdf; fix #8 clarify what wild card search is

Of the non-machine readable types, the most noteable is the Portable
Document Format (PDF)[^pdf]. These can be broken out into two groups: text based
PDFs and scanned PDFs. The former are converted from digital versions of
various kinds (MS Word, OpenOffice, LaTeX, markdown, etc.), while the
latter are created by scanning print articles to a PDF format. Text-based
PDFs are much better for text mining purposes as plain text can be extracted
easily in R with [pdftools][], a binding to [libpoppler][]. However, with
scanned PDFs, text must be extracted using Optimal Character Recognition
(OCR; see R package [tesseract][]), which isn't always a clean solution,
especially compared to true text based PDFs.

@@

On top of the vast array of different data sources is the varied ways that 
search is implemented in each source. Most sources are probably using Solr
or Elasticsearch under the hood, though we can't know this for sure as most 
do not make their software infrastructure public knowledge. Nonetheless, data
sources differ in how search works from the user perspective. For example, some 
provide wild card search and some do not. Some sources are searching full text
of articles, while others only search metadata (i.e., title, authors, abstract).
In addition, each source has a different set of metadata/full text available.
In brief, the same search against different sources produces different results.
Some text mining research articles perform the same search against many
different sources (refs), while others choose just one source.

>>>>

Of the non-machine readable types, the most noteable is the Portable
Document Format (PDF)[^pdf]. These can be broken out into two groups: text based
PDFs and scanned PDFs (images of text). The former are converted from digital
versions of various kinds (MS Word, OpenOffice, LaTeX, markdown, etc.), while the
latter are created by scanning print articles to a PDF format. Text-based
PDFs are much better for text mining purposes as plain text can be extracted
easily in R with [pdftools][], a binding to [libpoppler][]. However, with
scanned PDFs, text must be extracted using Optimal Character Recognition
(OCR; see R package [tesseract][]), which isn't always a clean solution,
especially compared to true text based PDFs.

@@

On top of the vast array of different data sources is the varied ways that 
search is implemented in each source. Most sources are probably using Solr
or Elasticsearch under the hood, though we can't know this for sure as most 
do not make their software infrastructure public knowledge. Nonetheless, data
sources differ in how search works from the user perspective. For example, some 
provide wild card/fuzzy search (i.e., 'appl\*' includes results for 'apple',
'application', etc.) and some do not. Some sources are searching full text
of articles, while others only search metadata (i.e., title, authors, abstract).
In addition, each source has a different set of metadata/full text available.
In brief, the same search against different sources produces different results.
Some text mining research articles perform the same search against many
different sources (refs), while others choose just one source.

### 2019-08-23 manuscript.tex
fix #7 clarify scanned pdf; fix #8 clarify what wild card search is

Of the non-machine readable types, the most noteable is the Portable
Document Format (PDF)\footnote{\url{https://en.wikipedia.org/wiki/PDF}}.
These can be broken out into two groups: text based PDFs and scanned
PDFs. The former are converted from digital versions of various kinds
(MS Word, OpenOffice, LaTeX, markdown, etc.), while the latter are
created by scanning print articles to a PDF format. Text-based PDFs are
much better for text mining purposes as plain text can be extracted
easily in R with \href{https://github.com/ropensci/pdftools}{pdftools},
a binding to \href{https://poppler.freedesktop.org/}{libpoppler}.
However, with scanned PDFs, text must be extracted using Optimal
Character Recognition (OCR; see R package
\href{https://github.com/ropensci/tesseract}{tesseract}), which isn't
always a clean solution, especially compared to true text based PDFs.

@@

On top of the vast array of different data sources is the varied ways
that search is implemented in each source. Most sources are probably
using Solr or Elasticsearch under the hood, though we can't know this
for sure as most do not make their software infrastructure public
knowledge. Nonetheless, data sources differ in how search works from the
user perspective. For example, some provide wild card search and some do
not. Some sources are searching full text of articles, while others only
search metadata (i.e., title, authors, abstract). In addition, each
source has a different set of metadata/full text available. In brief,
the same search against different sources produces different results.
Some text mining research articles perform the same search against many
different sources (refs), while others choose just one source.

>>>>

Of the non-machine readable types, the most noteable is the Portable
Document Format (PDF)\footnote{\url{https://en.wikipedia.org/wiki/PDF}}.
These can be broken out into two groups: text based PDFs and scanned
PDFs (images of text). The former are converted from digital versions of
various kinds (MS Word, OpenOffice, LaTeX, markdown, etc.), while the
latter are created by scanning print articles to a PDF format.
Text-based PDFs are much better for text mining purposes as plain text
can be extracted easily in R with
\href{https://github.com/ropensci/pdftools}{pdftools}, a binding to
\href{https://poppler.freedesktop.org/}{libpoppler}. However, with
scanned PDFs, text must be extracted using Optimal Character Recognition
(OCR; see R package
\href{https://github.com/ropensci/tesseract}{tesseract}), which isn't
always a clean solution, especially compared to true text based PDFs.

@@

On top of the vast array of different data sources is the varied ways
that search is implemented in each source. Most sources are probably
using Solr or Elasticsearch under the hood, though we can't know this
for sure as most do not make their software infrastructure public
knowledge. Nonetheless, data sources differ in how search works from the
user perspective. For example, some provide wild card/fuzzy search
(i.e., `appl*' includes results for `apple', `application', etc.) and
some do not. Some sources are searching full text of articles, while
others only search metadata (i.e., title, authors, abstract). In
addition, each source has a different set of metadata/full text
available. In brief, the same search against different sources produces
different results. Some text mining research articles perform the same
search against many different sources (refs), while others choose just
one source.
